---
title: "Untitled"
author: "Ruth Muriithi"
date: "1/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Reading the Data
```{r}
#Installing Packages 
#

suppressPackageStartupMessages(install.packages("data.table"))
require(data.table)

suppressPackageStartupMessages(install.packages("anomalize"))
library(anomalize)

suppressPackageStartupMessages(install.packages("dplyr"))
require(dplyr)

suppressPackageStartupMessages(install.packages("tibbletime"))
require(tibbletime)

suppressPackageStartupMessages(install.packages("tibble"))
require(tibble)

#Loading our data 
url = "http://bit.ly/CarreFourSalesDataset"
df.cf <- read.csv(url)

```

Checking the Data 

```{r}
#Previewing our dataset 
#
head(df.cf)

str(df.cf)
```

Basic data preview

```{r}
#Preview the column names 
#

colnames(df.cf)

#shape of the dataset 
#

dim(df.cf)

```
Tidying the Data

```{r}


#Checking for duplicates 
#

anyDuplicated(df.cf)

#Checking for missing values 
#

colSums(is.na(df.cf))


```

Conversion to date data type

```{r}

df.cf$Date <- as.Date(df.cf$Date, format = "%m/%d/%y")
str(df.cf)

#Convertion to POCIXct type
df.cf$Date <- as.POSIXct(df.cf$Date)


df.cf <- df.cf %>% as.tibble()
```

# Implementing the Solution 

```{r}
#Plotting anomalies
df.cf %>%
  time_decompose(Sales, method = "stl",  frequency = "auto", trend = "auto") %>%  
  anomalize(remainder, method = "gesd", alpha = 0.05, max_anoms = 0.1) %>% 
  plot_anomaly_decomposition()

```

#Recomposing

```{r}

df.cf %>% 
  time_decompose(Sales, method = "stl",  frequency = "auto", trend = "auto") %>% 
  anomalize(remainder,  method = "gesd", alpha = 0.05, max_anoms = 0.1) %>%
  time_recompose() %>%  
  plot_anomalies(time_recomposed = TRUE, ncol = 3, alpha_dots = 0.5)
```

# finding anomalies


```{r}
anomalies = df.cf %>% 
  time_decompose(Sales, method = "stl",  frequency = "auto", trend = "auto") %>%  
  anomalize(remainder,  method = "gesd", alpha = 0.05, max_anoms = 0.1) %>%  
  time_recompose() %>%  
  filter(anomaly == 'Yes')
anomalies
```

The dates listed above have anomalies.

# Follow up Questions 

1. Did we have the right data? **Yes we did**

2. Did we have the right question? **Yes we did**

3. Do we need to do anything else to answer the question? **No**

# Conclusion and Recommendation

I recommend that the carrefour marketing team keenly looks at the anomalies identified, and not be quick to assume they part of the trend while coming up with the strategy. 