---
title: "Market Basket Analysis"
author: "Ruth Muriithi"
date: "1/23/2021"
output: html_document
---

In this session, we will go through an example of association rules using the arules package. The documentation of this package can be found by visiting the following link:
*https://www.rdocumentation.org/packages/arules/versions/1.6-4.*

Below is an extract from its documentation:

"It provides the infrastructure for representing, manipulating and analyzing transaction data and patterns (frequent itemsets and association rules). It also provides interfaces to C implementations of the association mining algorithms Apriori and Eclat."$

**Defining the metric for success**

The project will be considered a success if we are able to build an apriori model. 

Reading the Data
```{r}
#Installing packages

#install.packages("arules")
library(arules)

#Loading our data 
#http://bit.ly/SupermarketDatasetII
#

url = "http://bit.ly/SupermarketDatasetII"
Transactions<-read.transactions(url, sep = ",")
Transactions
```
Checking the Data 

```{r}
# Verifying the object's class
# ---
# This should show us transactions as the type of data that we will need
# ---
# 
class(Transactions)
```
Previewing our first 5 transactions

```{r}
inspect(Transactions[1:5])

```
Previewing the items that make up our dataset,

```{r}
items<-as.data.frame(itemLabels(Transactions))
colnames(items) <- "Item"
head(items, 10)
```
 
Implementing the Solution

```{r}
# Generating a summary of the transaction dataset
#Informing the marketing department on the most relevant marketing strategies that will result in the highest no. of sales
# 
summary(Transactions)

```
As you can see above the items that are frequently purchased include: 
>
- Mineral Water 
- Eggs 
- Spaghetti 
- French Fries 
- Chocolate 


Exploring the frequency of some articles 

```{r}

itemFrequency(Transactions[, 5:10],type = "absolute")
round(itemFrequency(Transactions[, 5:10],type = "relative")*100,2)
```
Producing a chart of frequencies and fitering and Displaying top 10 most common items in the transactions dataset and the items whose relative importance is at least 10%

```{r}
par(mfrow = c(1, 2))
# plot the frequency of items

itemFrequencyPlot(Transactions, topN = 10,col="darkgreen")
itemFrequencyPlot(Transactions, support = 0.1,col="darkred")

```
Building a model based on association rules we use Min Support as 0.001 and confidence as 0.8

```{r}
rules <- apriori (Transactions, parameter = list(supp = 0.001, conf = 0.8))
rules
```
 
Building a apriori model with Min Support as 0.002 and confidence as 0.8

```{r}
rules2 <- apriori (Transactions,parameter = list(supp = 0.002, conf = 0.8)) 

```

Building apriori model with Min Support as 0.002 and confidence as 0.6

```{r}
rules3 <- apriori (Transactions, parameter = list(supp = 0.001, conf = 0.6)) 

rules2

rules3
```

Observing rules built in our model i.e. first 5 model rules
 

```{r}
inspect(rules[1:5])

# Ordering these rules by a criteria 
# 

rules<-sort(rules, by="confidence", decreasing=TRUE)
inspect(rules[1:5])
```

# Follow up Questions

1. Did we have the right data? **Yes we did**

2. Did we have the right question? **Yes we did**

3. Do we need to do anything else to answer the question? **No**


# Conclusion and Recommendation

As you can see above the items that are frequently purchased include:

>Mineral Water
Eggs
Spaghetti
French Fries
Chocolate

$Strategy$ 

1. Ensure that stock for the above items is always sufficient. 
2. strategically place the above items: place them where they can easily be seen. This may result to a client purchasing some of these items despite it not being on the shopping list hence higher sales.
3, Curate favourable offers