---
title: "PCA + Feature Selection"
author: "Ruth Muriithi"
date: "1/23/2021"
output: html_document
---

#Specifying the Question

*Undertaking a project on behalf of Carrefour Kenya, that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales.*

#Defining the metric for success

The ability adequately selecting our features and reduce our dimensions is what will determine whether the project is a success or not.

#Understanding the context

Carrefour is a multinational corporation of French origin specialized in retail. I was founded in 1958, making it 62 years old.

In 2016 Carrefour set up shop in Kenya, It is currently in 15 countries and has over 250 hypermarkets.

This project is being done on behalf of Carrefour Kenya, I will be undertaking a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax).

#Recording the experimental design
The findings of my analysis will be recorded in this markdown file and uploaded on github.

#Data reading and checking

```{r}
df <- read.csv("http://bit.ly/CarreFourDataset")
head(df)

```

```{r}
summary(df)
```
```{r}
#Checking the data type of each column 
#

str(df)
```

*Data Cleaning*

```{r}
#Preview the column names 
#
df.copy <- df
colnames(df)
```


```{r}
#shape of the dataset 
#

dim(df)
```


```{r}
# Tidying the Data

#Checking for duplicates 
#

anyDuplicated(df)
```
There are no duplicates in our dataset.

```{r}
#Checking for missing values 
#

colSums(is.na(df))
```
None of the columns have a missing value.

We'll factorise our character 

```{r}
## Encoding 
df$Branch <- as.factor(df$Branch)
df$Product.line <- as.factor(df$Product.line)
df$Payment <- as.factor(df$Payment)
df$Customer.type <- as.factor(df$Customer.type)
df$Gender <- as.factor(df$Gender)

str(df)

```

```{r}
#Label Encoding 
#Customer_type 
#
df$Customer.type <- ifelse(df$Customer.type == "Member",1,0)
df$Gender <- ifelse(df$Gender == "Female",1,0)

table(df$Customer.type)
table(df$Gender)
```
```{r}
head(df)
```

```{r}
#install.packages("caret")
library(caret)
require(caret)
df <- df[-c(1,9,10)]
dmy <- dummyVars(" ~ .", data = df, fullRank = T)
df.final <- data.frame(predict(dmy, newdata = df))
df.final
```

# Exploratory Analysis

## Univariate Analysis 
```{r}
#Summary statistics of our dataset

summary(df.copy)
```

## Bivariate Analysis


```{r}
# Create correlation matrix of data
df.num <- Filter(is.numeric, df.copy)
df.num <- df.num[,c(1:4, 6, 7,8)] 
res <- cor(df.num)
round(res, 4)
```
We dropped the gross margin percentage column because it has a standard deviation of zero.


# Multivariate Analysis
```{r}
head(df.final)
```

### PCA
```{r}
#Dropping the gross_margin_percentage
df.final <- df.final[c(1:15, 17,18,19)]
df.pr <- prcomp(df.final, center = TRUE, scale = TRUE)
summary(df.pr)

screeplot(df.pr, type = "l", npcs = 15, main = "Screeplot of the first 10 PCs")
abline(h = 1, col="red", lty=5)
legend("topright", legend=c("Eigenvalue = 1"),
       col=c("red"), lty=5, cex=0.6)
cumpro <- cumsum(df.pr$sdev^2 / sum(df.pr$sdev^2))
plot(cumpro[0:15], xlab = "PC #", ylab = "Amount of explained variance", main = "Cumulative variance plot")
abline(v = 6, col="blue", lty=5)
abline(h = 0.88759, col="blue", lty=5)
legend("topleft", legend=c("Cut-off @ PC6"),
       col=c("blue"), lty=5, cex=0.6)

```

#Claculating the Eigenvalue per component
#It is advised to only keep components with an eigenvalue greater than 1

```{r}
df.pr$sdev ^ 2
```



The first 9 variables have a eigenvalue greater than 1, We shall use them for our modelling. 

```{r}

print(df.pr$rotation)
```

#A biplot, it will help us see the grouping of different variables


```{r}
library(tidyselect)
library(tidyverse)
df.pr %>% biplot(cex = .5)
```


# Implementing the Solution 

### Feature Ranking
```{r}
#suppressWarnings(suppressMessages(if(!require(FSelector, quietly=TRUE)) #install.packages("FSelector")))
library(FSelector)


#Ranking variables

Scores <- linear.correlation(df.final)
Scores %>% arrange(desc(Scores$attr_importance))

```
The list of important features in descending order is as shown above. 

The table below shows the top 5 features.

```{r}
Subset <- cutoff.k(Scores, 5)
as.data.frame(Subset)
```
As you can see our top five performing features are:
>1. Branch.C			
2. Rating			
3. Product.line.Food.and.beverages			
4. Product.line.Sports.and.travel			
5. Payment.Credit.card

# Follow up Questions 

1. Did we have the right data? **Yes we did**

2. Did we have the right question? **Yes we did**

3. Do we need to do anything else to answer the question? **No**

# Conclusion and Recommendation

If we were to use five features to build our model, we would use the above. 

